# R-INLA modeling dry wells
# Author: José M. Rodríguez-Flores
# Last version: 06/05/2023
# Repo: 

```{r}
library(tidyverse)       
library(patchwork)        
library(ggtext)           
library(scales)          
library(modelr)           
library(distributional)   
library(kableExtra)       
library(INLA)
library(INLAutils)
library(brinla)
library(usdm)
library(lares)
library(PresenceAbsence)
library(pROC)
library(geosphere)
library(ggdist)
library(PresenceAbsence)
library(psych)
library(sf)
library(blockCV)
library(spatialRF)
library(vtable)
library(gstat)
library(lme4)
library(reshape2)
library(ggcorrplot)
library(ggnewscale)
library(ggside)
```


```{r}
# Basins Study Area
shp_gwb <- st_read('./Study_Area/sa_basins.shp') 
# convert to sf object
gwb_p <- st_as_sf(shp_gwb,
       remove = F, 
       crs = 4326) 
# define projection
gwb_p <- gwb_p %>% st_transform(crs = 3578) # meters better than long lat for R-INLA
```

```{r}
data_model_5mi <- readRDS("./Data_model/wells_data_5mi.rds")
data_model_3mi <- readRDS("./Data_model/wells_data_3mi.rds")
```

# Transform coords to UTM - meters

```{r}
data_model_3mi <- st_as_sf(data_model_3mi,
       remove = FALSE, 
       coords=c('lon','lat'),
       crs = 4326) |> st_transform(crs = 3578)
data_model_5mi <- st_as_sf(data_model_5mi,
       remove = FALSE, 
       coords=c('lon','lat'),
       crs = 4326) |> st_transform(crs = 3578)

data_model_3mi <- data_model_3mi |> mutate(x = sf::st_coordinates(geometry)[,1],
                                           y = sf::st_coordinates(geometry)[,2])
data_model_5mi <- data_model_5mi |> mutate(x = sf::st_coordinates(geometry)[,1],
                                           y = sf::st_coordinates(geometry)[,2])

data_model_3mi <- as_tibble(data_model_3mi) |> dplyr::select(-geometry)
data_model_5mi <- as_tibble(data_model_5mi) |> dplyr::select(-geometry)
```


# # Spatial autocorrelation
# Moran's autocorr plot

```{r}
predictors <- c("r_area_perennial","r_area_forage","r_area_annual","r_well_depth_gw_level","dom_wells_density","ag_wells_density","Poverty","Surface_Water")

Loc <- data_model_3mi[, c("lon", "lat")] #Spatial locations
Distances <- distm(Loc)/1000 # Km
Loc["x"] <- Loc$lon
Loc["y"] <- Loc$lat

#distance thresholds (same units as distance_matrix)
distance.thresholds <- c(5, 20, 35, 50, 100)

spatialRF::plot_training_df_moran(
  data = data_model_3mi,
  dependent.variable.name = "dry_condition",
  predictor.variable.names = predictors,
  distance.matrix = Distances,
  distance.thresholds = distance.thresholds,
  fill.color = viridis::viridis(
    100,
    option = "C",
    direction = -1),
  point.color = "black") +
  labs(x = "Distance threshold (km)")
```
#
# Correlations
#
```{r}
cor <- data_model_3mi |>
  dplyr::select(r_area_perennial,
                r_area_annual,
                r_area_forage,
                r_well_depth_gw_level,
                ag_wells_density,
                dom_wells_density,
                Poverty,
                Surface_Water,
                dry_condition)
```
#
```{r}
# Function to get asteriks
labs.function = function(x){
  case_when(x >= 0.05 ~ "",
            x < 0.05 & x >= 0.01 ~ "*",
            x < 0.01 & x >= 0.001 ~ "**",
            x < 0.001 ~ "***")
}


cor$Surface_Water[is.na(cor$Surface_Water)] = 0
cor_m <- round(cor(cor),1)
p.mat <- cor_pmat(cor)
p.labs = as.data.frame(p.mat)  %>%
  mutate_all(labs.function)

p.labs$Var1 = as.factor(rownames(p.labs))
p.labs = melt(p.labs, id.vars = "Var1", variable.name = "Var2", value.name = "lab")


cor.plot <- ggcorrplot(cor_m,hc.order=TRUE,type="lower",colors = c("darkblue", "white", "red"),lab = TRUE) +  theme_bw(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="",y="")

p.labs$in.df = ifelse(is.na(match(paste0(p.labs$Var1, p.labs$Var2),
                                  paste0(cor.plot[["data"]]$Var1, cor.plot[["data"]]$Var2))),
                      "No", "Yes")

p.labs = dplyr::select(p.labs |> filter(in.df == "Yes"), -in.df)

# Add asteriks to ggcorrplot
cor.plot.labs = cor.plot +
  geom_text(aes(x = p.labs$Var1,
                y = p.labs$Var2),
            label = p.labs$lab,
            nudge_y = 0.25,
            size = 5) +
  theme(panel.grid.major = element_blank())

cor.plot.labs
```


```{r}
# Variance Inflation Factor (VIF) values
vif(as.data.frame(cor))
```

# Distances among wells

```{r}
Loc2 <- data_model_3mi[, c("x", "y")] #Spatial locations
Distances <- dist(Loc2)*0.000621371 #meters * miles
hist(Distances,xlab="Distances (mi)")
summary(c(Distances))
```

# Functions

## Function for data splitting training and validation (cross-validation)

```{r}
split_data <- function(data){
# Split Training and validation per data frame
pa_data <- sf::st_as_sf(data |> dplyr::select(lon,lat,dry_condition), coords = c("lon", "lat"), crs = 4326)
sb1 <- cv_spatial(x = pa_data,
                  column = "dry_condition", # the response column (binary 1/0)
                  k = 4, # number of folds
                  size = 1600, # size of the blocks in meters (1 mi)
                  selection = "random", # random blocks-to-fold
                  iteration = 50,
                  seed = 30) # find evenly dispersed folds

folds <- sb1$folds_list
training_data <- data[as.vector(folds[[1]][[1]]),]
testing_data <- data[as.vector(folds[[1]][[2]]),]
return(list(training_data,testing_data))}
```

## Functions that we'll be using for summaries and visualizations

```{r}
random_fix_effects <- function(r){

fix_eff = tibble(names = names(r$marginals.fixed), m.fixed = r$marginals.fixed)
fix_eff = fix_eff |> 
  mutate(samples = purrr::map(m.fixed,~inla.smarginal(.x))) |>
  mutate(x = purrr::map(samples,~.x[['x']]))

df_fix = fix_eff |> dplyr::select(names,x) |> unnest(x)

func_odds <- function(x){exp(x)/(1 + exp(x))}

df_fix <- df_fix|> mutate(odds=sapply(x,func_odds))

# Samples distribution random effects
random_eff = tibble('names' = names(r$marginals.random),
                    'm.random' = r$marginals.random)

random_eff = random_eff |> 
  mutate(index = purrr::map(m.random,~names(.x))) |> unnest(index)

random_eff = random_eff |>
  mutate(samples = purrr::map2(m.random,index,~inla.smarginal(.x[[.y]]))) |>
  mutate(x = purrr::map(samples,~.x[['x']]))

df_random = random_eff |> dplyr::select(names,index,x) |> unnest(x)

func_odds <- function(x){exp(x)/(1 + exp(x))}

df_random <- df_random |> mutate(odds=sapply(x,func_odds))
return(list(df_fix,df_random))}

# Plots random and fixed effects
plot_dist_f_eff <- function(name,df){
  plot <- df |> filter(names==name) |>  
  ggplot(aes(x = x)) +
  stat_slab(aes(
    fill = after_stat(cut_cdf_qi(cdf,
                                 .width = c(.5, .75, .95),
                                 labels = percent_format(accuracy = 1)))),
    position='dodge', point_interval = mean_qi)  +
     stat_pointinterval(.width = c(.5, .75, .95),show.legend = c(size = F),point_size=3) +
  scale_size_continuous(range = c(0.5, 10)) +
  scale_fill_manual(values = c("#2c7fb8","#41b6c4","#a1dab4"),
                    na.value = NA,
                    na.translate = FALSE) +
  labs(fill="Quantile",y="",x="",title=name) + geom_vline(xintercept=0,color="red") +
    theme_bw(base_size = 14)+
  theme(panel.background = element_blank(),
        plot.title = element_text(size=9),
        panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()) + 
  scale_y_continuous(expand = c(0, 0))
  
  return(plot)}

plot_dist_f_eff2 <- function(df){
  plot <- df |>  
  ggplot(aes(x = x,y=names)) +
  stat_slab(aes(
    fill = after_stat(cut_cdf_qi(cdf,
                                 .width = c(.5, .75, .95),
                                 labels = percent_format(accuracy = 1)))),
    position='dodge', point_interval = mean_qi) +
     stat_pointinterval(.width = c(.5, .75, .95),show.legend = c(size = F),point_size=3) +
  scale_size_continuous(range = c(0.5, 10)) +
   scale_fill_manual(values = c("#2c7fb8","#41b6c4","#a1dab4"),
                     na.value = NA,
                     na.translate = FALSE) +
  labs(fill="Interval",y="",x="") + geom_vline(xintercept=0,color="red") +
     theme_bw(base_size = 14)+
  theme(panel.background = element_blank(),
        plot.title = element_text(size=9),
        panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(0, 0, 0, 0, "pt")) 
  
  return(plot)}

plot_dist_r_eff <- function(name,df){
  plot <- df |> filter(names==name) |>  
  ggplot(aes(x = x,y=index)) +
  stat_slab(aes(
    fill = after_stat(cut_cdf_qi(cdf,
                                 .width = c(.5, .75, .95),
                                 labels = percent_format(accuracy = 1)))),
    position='dodge', point_interval = mean_qi)  +
     stat_pointinterval(.width = c(.5, .75, .95),show.legend = c(size = F),point_size=3) +
  scale_size_continuous(range = c(1, 11)) +
    scale_fill_manual(values = c("#2c7fb8","#41b6c4","#a1dab4"),
                      na.value = NA,
                      na.translate = FALSE) +
  labs(fill="Interval",y="",x="",title="") + geom_vline(xintercept=0,color="red") +
     theme_bw(base_size = 14)+
  theme(panel.background = element_blank(),
        plot.title = element_text(size=9),
        panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(0, 0, 0, 0, "pt")) 
  
  return(plot)}
```

## INLA Functions

```{r}
# Define stack with inputs for INLA
build_stack <- function(data_model){
# Coordinates
coo <- cbind(data_model$x, data_model$y) # Coordinates - meters UTM
# Create the mesh
RangeGuess <- 80000 # meters
MaxEdge  <- RangeGuess/5
ConvHull <- inla.nonconvex.hull(coo,convex=-0.08)		
# convex factor changes the inner boundary to be closer to points
# the non-convex hull approach: we might want to use a mesh which is based on a non-convex hull to avoid adding many small triangles outside the domain of interest (more triangles = larger computation times)
mesh    <- inla.mesh.2d(loc = coo, #location coordinates that are used as initial mesh vertices
                        boundary = ConvHull,
                        max.edge = c(1, 5) * MaxEdge, #values denoting the maximum allowed triangle edge lengths in the region and in the extension outside
                        cutoff  = MaxEdge / 5) #minimum allowed distance between points  
# Define the weight factors or projector matrix

plot(mesh)
points(coo, col = "red")
plot(gwb_p,add=T,border='darkgreen',color=NA,lwd=2)

# Projection
A <- inla.spde.make.A(mesh = mesh, loc = coo)

# Create SPDE function (w matern covariance)
spde <- inla.spde2.pcmatern(mesh = mesh,
                          # alpha = 2,
                          # constr=TRUE,
                          prior.range = c(64373,0.01),
                          prior.sigma =  c(0.01,0.001))



# Index spatial field
indexs <- inla.spde.make.index(name = "field",
                               n.spde = spde$n.spde)

print(lengths(indexs))

stackfit <- inla.stack(
  tag = "fit",
  data = list(y = data_model$dry_condition),
  A = list(1, A),
  effects = list(data.frame(intercept = rep(1,nrow(data_model)),
                            r_area_perennial = data_model$r_area_peren_sd,
                            r_area_annual = data_model$r_area_annual_sd,
                            r_area_forage = data_model$r_area_for_sd,
                            r_well_depth_gw_level  = data_model$r_welldepth_gwdepth,
                            surface_water  = data_model$sw_sd,
                            ag_wells_density = data_model$ag_wells_density_sd,
                            dom_wells_density = data_model$dom_wells_density_sd,
                            Year_index = data_model$Year_index,
                            basin_index = data_model$basin_index,
                            poverty = data_model$Poverty_sd),
                            field = indexs)
)

return(list(stackfit,spde,mesh))}
```

```{r}
stack_validation <- function(data,mesh,field.spde){
indexs <- inla.spde.make.index(name = "field",
                               n.spde = field.spde$n.spde)
LocPred <- cbind(data$x, data$y)
A.pred <- inla.spde.make.A(mesh = mesh, loc = LocPred)
N <- nrow(data)
stack <- inla.stack(
  tag = "validation",
  data = list(y = data$dry_condition),  
  A = list(1, A.pred),
  effects = list(data.frame(intercept = rep(1,nrow(data)),
                            r_area_perennial = data$r_area_peren_sd,
                            r_area_annual = data$r_area_annual_sd,
                            r_area_forage = data$r_area_for_sd,
                            r_well_depth_gw_level  = data$r_welldepth_gwdepth,
                            surface_water  = data$sw_sd,
                            ag_wells_density = data$ag_wells_density_sd,
                            dom_wells_density = data$dom_wells_density_sd,
                            Year_index = data$Year_index,
                            basin_index = data$basin_index,
                            poverty = data$Poverty_sd),
                            field = indexs)
)

return(stack)}
```

# INLA Model Calibration function
```{r}
INLA_model <- function(model,buffer){
  
if(buffer=="3mi"){
  stack1 = stack_3mi[[1]]
  spde1 = stack_3mi[[2]]
}else if(buffer=="5mi"){
  stack1 = stack_5mi[[1]]
  spde1 = stack_5mi[[2]]
}
  
formula <- tidy_models |> filter(buffer_size == buffer & names_model == model) |> pull(models)

res <- INLA::inla(formula[[1]],
  family = "binomial",
  Ntrials = 1,
  control.family = list(link = "logit"),
  data = inla.stack.data(stack1,spde=spde1),
  verbose=T,
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE,config=TRUE,residuals=TRUE), 
  control.predictor = list(
    compute = TRUE, link = 1,
    A = inla.stack.A(stack1)
  )
)

return(list(res))}
```

# INLA Model Prediction function
```{r}
INLA_pred_model <- function(model,buffer){
  
if(buffer=="3mi"){
  stack1 = stack_pred_3mi
  spde1 = stack_3mi[[2]]
}else if(buffer=="5mi"){
  stack1 = stack_pred_5mi
  spde1 = stack_5mi[[2]]
}
  
formula <- tidy_models |> filter(buffer_size == buffer & names_model == model) |> pull(models)

pred <- INLA::inla(formula[[1]],
  family = "binomial",
  Ntrials = 1,
  control.family = list(link = "logit"),
  data = inla.stack.data(stack1,spde=spde1),
  control.compute = list(cpo = TRUE,config=TRUE,residuals=TRUE), 
  control.predictor = list(
    compute = TRUE, link = 1,
    A = inla.stack.A(stack1)
  )
)

return(list(pred))}
```

# Performance function
```{r}
performance_stats <- function(buffer,result){
  
if(buffer=="3mi"){
data_model = data_3mi[[1]]
}else if(buffer=="5mi"){
data_model = data_5mi[[1]]
}
  
index.fit <- inla.stack.index(stack_3mi[[1]], tag = "fit")$data
DATA <- data.frame(
  ID = c(1:nrow(data_model)),
  observed = data_model$dry_condition,
  predicted = result$summary.fitted.values[index.fit,"mean"])  

AUC <- PresenceAbsence::auc(DATA, st.dev = TRUE, which.model = 1, na.rm = FALSE)
DIC <- result$dic$dic
WAIC <- result$waic$waic
CPO.SCORE <- -sum(log(result$cpo$cpo))
auc.roc.plot(DATA, threshold=101, which.model=1, model.names=c("Best model"),
             na.rm=TRUE, xlab="1-Specificity (false positives)", ylab="Sensitivity (true positives)", main="ROC Plot", color=TRUE, line.type=TRUE, lwd=1, mark=0, mark.numbers=TRUE, opt.thresholds=TRUE)
AUC.ROC.PLOT = recordPlot() 
conf_matrix <- cmx(DATA, threshold = 0.5, which.model = 1, na.rm = FALSE)
return(list(AUC,DIC,WAIC,CPO.SCORE,AUC.ROC.PLOT,as.data.frame(conf_matrix)))
}
```

# Modeling for results shown in manuscript

```{r}
# Stacks for model calibration
data_3mi <- split_data(data_model_3mi)
data_5mi <- split_data(data_model_5mi)
```

# Let's do it tidy 

```{r}
m1 <- y ~ -1 + intercept + f(field, model = spde) + r_area_perennial +  r_area_annual +  r_area_forage

m2 <- y ~ -1 + intercept + f(field, model = spde) + surface_water + ag_wells_density + dom_wells_density + r_well_depth_gw_level

m3 <-  y ~ -1 + intercept + f(field, model = spde) + r_area_perennial +  r_area_annual +  r_area_forage +  r_well_depth_gw_level + surface_water + dom_wells_density  + ag_wells_density 

m5 <-   y ~ -1 + intercept + f(field, model = spde) + r_area_perennial +  r_area_annual +  r_area_forage +   r_well_depth_gw_level + surface_water + dom_wells_density + ag_wells_density +  f(basin_index, model = "iid") + ag_wells_density:r_area_perennial + ag_wells_density:r_area_annual + ag_wells_density:r_area_forage 

m6 <- y ~ -1 + intercept + f(field, model = spde) + r_area_perennial +  r_area_annual +  r_area_forage + r_well_depth_gw_level + surface_water + dom_wells_density + ag_wells_density +  f(basin_index, model = "iid") + ag_wells_density:r_area_perennial + ag_wells_density:r_area_annual + ag_wells_density:r_area_forage + poverty 

m4 <- y ~ -1 + intercept + f(field, model = spde) + r_area_perennial +  r_area_annual +  r_area_forage +   surface_water + dom_wells_density + ag_wells_density +  f(basin_index, model = "iid") + ag_wells_density:r_area_perennial + ag_wells_density:r_area_annual + ag_wells_density:r_area_forage + poverty
```

```{r}
list_models <- c(m1,m2,m3,m4,m5,m6)
names_models <- c("m1","m2","m3","m4","m5","m6")
```

```{r}
# Create tibble
tidy_models <- tibble(buffer_size = c("3mi","5mi"))

tidy_models <- tidy_models |> mutate(models = list(list_models)) |> unnest(models)

tidy_models <- tidy_models |> mutate(names_model = rep(names_models,2))
```

```{r}
# Stacks for model calibration
stack_3mi <- build_stack(data_3mi[[1]])
stack_5mi <- build_stack(data_5mi[[1]])
```

```{r}
# Run models
tidy_models <- tidy_models |> rowwise() |>
  mutate(results = purrr::map2(names_model,buffer_size,~INLA_model(.x,.y),progress=T))
```

```{r}
# Add statistics
tidy_models <- tidy_models |> rowwise() |> 
  mutate(stats = map2(buffer_size,results,~performance_stats(.x,.y))) |> ungroup()
```

# Performance Calibration
```{r}
results_models <- tidy_models |> dplyr::select(-results) |> 
  mutate(AUC = map_dbl(stats,~.x[[1]][["AUC"]]),
         AUC.SD = map_dbl(stats,~.x[[1]][["AUC.sd"]]),
         DIC = map_dbl(stats,~.x[[2]]),
         WAIC  = map_dbl(stats,~.x[[3]]),
         CPO.SCORE = map_dbl(stats,~.x[[4]]),
         AUC.ROC.PLOT = map(stats,~.x[[5]]),
         CONF.M = map(stats,~.x[[6]]))
  
```

# Visualize Results best model

```{r}
res_m <- tidy_models$results[tidy_models$buffer_size == "3mi" & tidy_models$names_model == "m6"][[1]][[1]]
```

```{r}
# Summary Results
res_m$summary.fixed
```

```{r}
# Visualize
results <- random_fix_effects(res_m)

results[[2]]$index[results[[2]]$names=="basin_index" & results[[2]]$index == "index.1"] <- "MADERA"  

results[[2]]$index[results[[2]]$names=="basin_index" & results[[2]]$index == "index.2"] <- "KINGS"  

results[[2]]$index[results[[2]]$names=="basin_index" & results[[2]]$index == "index.3"] <- "KAWEAH" 

results[[2]]$index[results[[2]]$names=="basin_index" & results[[2]]$index == "index.4"] <- "TULE" 
# Plots
fix_effects <- plot_dist_f_eff2(results[[1]]) 

plist <- map(list(unique(results[[2]]$names)[2]),~plot_dist_r_eff(.x,results[[2]]))

rand_effects <- patchwork::wrap_plots(plist,ncol=1,guides = "collect")+ labs(x="Log-Odds")
```

```{r}
(fix_effects/plot_spacer()/rand_effects) + 
  plot_layout(guides="collect",heights = c(16,-1.6,4))  & 
  theme(legend.position = 'bottom')  & 
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 12)) 
```


```{r}
# Marginal Posterior Distribution Hyperparameters
library(brinla)
p_hyper <- bri.hyperpar.plot(res_m, together = FALSE) 
ggplot() + geom_line(data=p_hyper,aes(x=x,y=y)) + facet_wrap(~parameter,scales="free") + labs(title="Marginal Posterior Distribution Hyperparameters",y="Density",x="")
```

# Spde results GMRF based on spde

```{r}
#looking at the spatial field and what it looks like
mesh = stack_3mi[[3]]
A1.grid <- inla.mesh.projector(mesh, dims = c(50, 50))
inla.mesh.project(A1.grid, res_m$summary.random$field) %>%
as.matrix() %>%
as.data.frame() %>%
bind_cols(
expand.grid(x = A1.grid$x, y = A1.grid$y)
) %>%
filter(!is.na(ID)) -> eta_spde

meanspde <- ggplot(st_as_sf(data_3mi[[1]],coords = c(lon='x',lat='y'),crs=3578)) +
  geom_tile(data = eta_spde, aes(x = x, y = y, fill = mean)) +
  scale_fill_viridis_c() +
  geom_sf(alpha=0.2,color='orange') + 
  geom_sf(data=gwb_p,alpha = 1, color="black", lwd=0.7,fill=NA) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      legend.key.size = unit(0.9, 'cm'),
      legend.text=element_text(size=7)) +
  labs(x="",y="")

sdspde <-ggplot(st_as_sf(data_3mi[[1]],coords = c(lon='x',lat='y'),crs=3578)) +
  geom_tile(data = eta_spde, aes(x = x, y = y, fill = sd)) +
  scale_fill_viridis_c() +
  geom_sf(alpha=0.2,color='orange') + 
  geom_sf(data=gwb_p,alpha = 1, color="black", lwd=0.7,fill=NA) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      legend.key.size = unit(0.9, 'cm'),
      legend.text=element_text(size=7)) +
  labs(x="",y="")

patchwork::wrap_plots(list(meanspde,sdspde),ncol=2)
```

# Residuals model

```{r}
## Extract residuals from binomial inla model
Index <- inla.stack.index(stack_3mi[[1]], tag  = "fit")$data 
RF1 <- data.frame(Residuals =  res_m$residuals$deviance.residuals, # Deviance residuals
                  Predictions = res_m$summary.fitted.values[Index,"mean"],
                  basin_index = data_3mi[[1]]$basin_index,
                  x = data_3mi[[1]]$x,
                  y = data_3mi[[1]]$y)

RF1 <- RF1 |> filter(!is.na(Residuals))


Loc <- RF1[, c("x", "y")] #Spatial locations
Distances <- dist(Loc)

point.color = viridis::viridis(100,option = "F")
line.color = "gray10"
fill.color = viridis::viridis(4,option = "F",alpha = 0.95)[2]

#getting residuals
residuals <- RF1$Residuals
residuals.df <- as.data.frame(residuals)

 #normality scores of the residuals
residuals.qq <- qqnorm(residuals, plot.it=FALSE) %>%
  as.data.frame()

#qqplot
p1 <- ggplot2::ggplot(data = residuals.qq) +
  ggplot2::geom_point(
    data = residuals.qq,
    ggplot2::aes(
      x = x,
      y = y,
      color = y
    )
  ) +
  ggplot2::scale_color_gradientn(colors = point.color) +
  ggplot2::stat_qq_line(
    ggplot2::aes(sample = residuals),
    col = line.color,
    linetype = "dashed"
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ylab("Residuals") +
  ggplot2::xlab("Theoretical")

#histogram
p2 <- ggplot2::ggplot(data = residuals.df) +
  ggplot2::aes(
    x = residuals
    ) +
  ggplot2::geom_histogram(
    # binwidth = bw,
    color = NA,
    fill = fill.color
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::geom_vline(
    xintercept = median(residuals),
    col = line.color,
    linetype = "dashed"
  ) +
  ggplot2::ylab("Count") +
  ggplot2::xlab("Residuals")

  #residuals vs predictions
p3 <- ggplot2::ggplot(RF1) +
    ggplot2::aes(
      x = Predictions,
      y = Residuals,
      color = Residuals
    ) +
    ggplot2::geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = line.color
    ) +
    ggplot2::geom_point() +
    ggplot2::scale_color_gradientn(colors = point.color) +
    ggplot2::theme_bw() +
    ggplot2::ggtitle("Residuals vs. predictions") +
    ggplot2::theme(
      plot.title = ggplot2::element_text(hjust = 0.5),
      legend.position = "none"
      )
variogram_data <- RF1
coordinates(variogram_data) <- ~x+y
proj4string(variogram_data)= sf::st_crs(3578)[[2]]
variogram_res <- gstat::variogram(Residuals~1, variogram_data, cressie = TRUE, cutoff = 100000)

p5 <- ggplot() +
  geom_point(data = variogram_res, aes(x = dist, y = gamma)) +
  geom_smooth(data = variogram_res, aes(x = dist, y = gamma),col = "red") +
  xlab("Distance m") + ylab("Semi-variance") +
  theme_bw() +
  theme(text = element_text(size = 11),
        legend.position="none") +
  ylim(0, 1) 

#final plot
residuals_plot <- ((p1 + p2) / p3) / (p5) + patchwork::plot_annotation(
      theme = ggplot2::theme(plot.title = element_text(hjust = 0.5))
  )
```

```{r}
residuals_plot
```

```{r}
# stop()
```

# Probability of dry well by variable plot

```{r}
#Function of Probability of occurrence curves for linear predictors 
plot_prob_occurrence <- function(data,var,name_var){
  
X.p <- data.frame(
  Intercept              = rep(1, nrow(data)),
  variable = data |> pull({{var}}),
  dry_condition = data$dry_condition)

MyData <- data.frame(variable = seq(from = min(X.p$variable),
                                            to = max(X.p$variable),
                                            length = 100))

Xmat <- model.matrix(~ variable,data = MyData)
Xmat <- as.data.frame(Xmat)
lcb <- inla.make.lincombs(Xmat)


M.p <- inla(dry_condition ~ variable,lincomb = lcb,
              control.predictor = list(link = 1,compute = TRUE),
              family = "binomial",
            Ntrials = 1,
            data = X.p)

Pred.pm <- M.p$summary.lincomb.derived[,c("mean","0.025quant", "0.975quant")]
# print(head(M.p$summary.lincomb.derived), digits = 2)
Pred.marg <- M.p$marginals.lincomb.derived
# Pred.marg[[1]] 

probability_f <- function(x) {exp(x) / (1 + exp(x))}

inla.qmarginal(c(0.025, 0.5, 0.975), 
               inla.tmarginal(probability_f , Pred.marg[[1]]))

inla.emarginal(probability_f, Pred.marg[[1]])

MyData$mu <- unlist(lapply(Pred.marg,function(x) inla.emarginal(probability_f, x)))
MyData$selo <- unlist(lapply(Pred.marg,function(x)inla.qmarginal(c(0.025),inla.tmarginal(probability_f, x))))
MyData$seup <- unlist(lapply(Pred.marg,function(x)inla.qmarginal(c(0.975),inla.tmarginal(probability_f, x))))

p <- X.p |> ggplot() +
  # geom_point(aes(y = dry_condition, x = variable),shape = 1,size = 1.5,alpha=0.15, colour="grey30") + 
   geom_dots(aes(x= variable, y =dry_condition,side = ifelse(dry_condition, "bottom","top")), 
            pch = 19, color = "grey35", scale = 0.2) +
  xlab(name_var) + ylab("") +
  theme_bw(base_size = 10) +
  geom_line(data = MyData,aes(x = variable,y = mu),colour = "darkorange",linewidth = 1.0) +
  geom_ribbon(data = MyData,aes(x = variable,ymax = seup,ymin = selo),alpha = 0.2) +
  theme(panel.grid.major  = element_blank(),
        panel.grid.minor  = element_blank())+scale_y_continuous(labels = label_percent())
  

return(p)
}
```

```{r}
forage_plot <- plot_prob_occurrence(data_model_3mi,r_area_forage,"Proportion Forage Crops")
perennials_plot <- plot_prob_occurrence(data_model_3mi,r_area_perennial,"Proportion Perennial Crops")
poverty_plot <- plot_prob_occurrence(data_model_3mi,Poverty,"Poverty")
annuals_plot <- plot_prob_occurrence(data_model_3mi,r_area_annual,"Proportion Annual Crops")
ratio_well_gw_plot <- plot_prob_occurrence(data_model_3mi,r_well_depth_gw_level,"Well depth/Groundwater level")
sd_sw_plot <- plot_prob_occurrence(data_model_3mi,sw_sd,"Standirized Surface Water Supply")
dom_wells_d_plot <- plot_prob_occurrence(data_model_3mi,dom_wells_density,"Domestic Wells Density")
ag_wells_d_plot <-plot_prob_occurrence(data_model_3mi,ag_wells_density,"Agricultural Wells Density")
```

```{r}
# Label
glob_lab <- "Probability of Domestic Well Failure"
p_lab <- 
  ggplot() + 
  annotate(geom = "text", x = 0.1, y = 0.1, label = glob_lab, angle = 90) +
  coord_cartesian(clip = "on")+
  theme_void()

plots <- ((forage_plot + perennials_plot + annuals_plot) / (dom_wells_d_plot + ag_wells_d_plot + ratio_well_gw_plot) /(poverty_plot + sd_sw_plot)) & plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 10))

(wrap_elements(p_lab) | wrap_elements(plots)) + plot_layout(widths = c(.1,5))
```

# Plot interaction effect
```{r}
interaction_plot <- function(cov_1,cov_2,name_1,name_2){
data = data_model_3mi
X.p <- data.frame(
  Intercept              = rep(1, nrow(data)),
  variable_1 = data |> dplyr::pull({{cov_1}}),
  variable_2 = data |> dplyr::pull({{cov_2}}),
  dry_condition = data$dry_condition)

comb_space <- raster(xmn=min(X.p$variable_1),
                     xmx=max(X.p$variable_1),
                     ymn=min(X.p$variable_2),
                     ymx=max(X.p$variable_2),
                     nrows=50, ncols=50)

MyData = as.data.frame(comb_space,xy=TRUE)
colnames(MyData)[1] <- "variable_1"
colnames(MyData)[2] <- "variable_2"

Xmat <- model.matrix(~variable_1+variable_2,data = MyData)
lcb <- inla.make.lincombs(as.data.frame(Xmat))


M.p <- inla(dry_condition ~ variable_1 + variable_2,
            lincomb = lcb,
              control.predictor = list(link = 1,compute = TRUE),
              family = "binomial",
            Ntrials = 1,
            data = X.p)

Pred.pm <- M.p$summary.lincomb.derived[,c("mean","0.025quant", "0.975quant")]
# print(head(M.p$summary.lincomb.derived), digits = 2)
Pred.marg <- M.p$marginals.lincomb.derived
# Pred.marg[[1]] 

probability_f <- function(x) {exp(x) / (1 + exp(x))}

inla.qmarginal(c(0.025, 0.5, 0.975), 
               inla.tmarginal(probability_f , Pred.marg[[1]]))

inla.emarginal(probability_f, Pred.marg[[1]])

MyData$mu <- unlist(lapply(Pred.marg,function(x) inla.emarginal(probability_f, x)))


p <- ggplot() +
  geom_raster(data= MyData,aes(fill = mu, x = variable_1, y = variable_2)) + 
labs(x=name_1,y=name_2,fill="Probability\nDry Well") +
  theme_bw(base_size=14) + scale_fill_viridis_c(option = "F",direction=-1,limits = c(0, 1)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_continuous(expand=c(0,0)) +
  new_scale_fill() 

p <- p +  geom_xsidedensity(data=X.p,aes(x=variable_1,y=stat(density)),color="darkorange2",size=1) +
  geom_ysidedensity(data=X.p,aes(y=variable_2,x=stat(density)),color="darkorange2",size=1) + theme_bw(base_size=11) +
  theme(ggside.axis.text.x=element_blank(),
        ggside.axis.ticks.x=element_blank(),
        ggside.axis.text.y=element_blank(),
        ggside.axis.ticks.y=element_blank()) 

return(p)
}
```

```{r}
perennial_ag_well_p <- interaction_plot(r_area_perennial,ag_wells_density,"Perennial crops Proportion","Agricultural Wells Density")
annual_ag_well_p <- interaction_plot(r_area_annual,ag_wells_density,"Annual Crops Proportion","Agricultural Wells Density")
forage_ag_well_p <- interaction_plot(r_area_forage,ag_wells_density,"Forage Crops Proportion","Agricultural Wells Density")

perennial_ag_well_p + annual_ag_well_p + forage_ag_well_p + plot_layout(guides = 'collect')
```


# Validation of selected model: Prediction validation data set

```{r}
# Frist get the stack from INLA for prediction
stack_pred_3mi <-  stack_validation(data_3mi[[2]],stack_3mi[[3]],stack_3mi[[2]])
stack_pred_5mi <-  stack_validation(data_5mi[[2]],stack_5mi[[3]],stack_5mi[[2]])
prediction_results <- INLA_pred_model ("m5","3mi")[[1]]
```

# Performance prediction for validation data set
```{r}
index.Pred <- inla.stack.index(stack_pred_5mi, tag = "validation")$data
DATA_valid <- data.frame(
  ID = c(1:nrow(data_3mi[[2]])),
  observed = data_3mi[[2]]$dry_condition,
  predicted = prediction_results$summary.fitted.values[index.Pred,"mean"])  

AUC_valid <- PresenceAbsence::auc(DATA_valid, st.dev = TRUE, which.model = 1, na.rm = FALSE)
CPO.SCORE <- -sum(log(prediction_results$cpo$cpo))
print(CPO.SCORE)
auc.roc.plot(DATA_valid, threshold=101, which.model=1, model.names=c("Best model"),
             na.rm=TRUE, xlab="1-Specificity (false positives)", ylab="Sensitivity (true positives)", main="ROC Plot - Validation Set", color=TRUE, line.type=TRUE, lwd=1, mark=0, mark.numbers=TRUE, opt.thresholds=TRUE)
AUC.ROC.PLOT_valid = recordPlot() 
conf_matrix_valid <- cmx(DATA_valid, threshold = 0.5, which.model = 1, na.rm = FALSE)
```

# Map prediction validation set

```{r}
library(ggspatial)
# Visualize where we were or not able to predict dry wells from the validation set
DATA_valid$x <- data_3mi[[2]]$x
DATA_valid$y <- data_3mi[[2]]$y

DATA_valid <- DATA_valid |> 
  mutate(Predicted = case_when(observed == 0 & predicted < 0.5 ~ "TRUE",
                               observed == 1 & predicted >= 0.5 ~ "TRUE",
                               .default = "FALSE"))

DATA_valid <- DATA_valid |> st_as_sf(coords=c("x","y"), crs = 3578) 

ggplot() + 
  geom_sf(data=gwb_p,color="black",fill=NA) +
  geom_sf(data=DATA_valid,aes(color=Predicted,shape=as.factor(observed)),
          size=2,
          alpha=0.6) +
  scale_color_manual(values=c("gray40","orange")) +
  labs(shape = "Dry Condition",
       color = "Predicted Corrrectly")+
  theme_bw(base_size = 14) +
  theme(axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      legend.position=c(0.3,0.2),
      legend.background = element_rect(fill='transparent'),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.border = element_blank())+
   annotation_north_arrow(
    location = "tl",
    pad_x = unit(4, "in"),
    pad_y = unit(0.1, "in"),
    style = north_arrow_fancy_orienteering
  )
```

# Test model without SPDE (test autocorrelation)

```{r}
logit_m6 <- glmer(dry_condition ~  r_area_peren_sd +  r_area_annual_sd +  r_area_for_sd + r_welldepth_gwdepth + sw_sd + dom_wells_density_sd +  r_area_peren_sd:ag_wells_density_sd +  r_area_annual_sd:ag_wells_density_sd +  r_area_for_sd:ag_wells_density_sd + Poverty_sd + (1|basin_index)+ ag_wells_density_sd, data = data_3mi[[1]], family = "binomial")

```

```{r}
summary(logit_m6)
```

```{r}
## Extract residuals from binomial inla model
RF2 <- data.frame(Residuals =  residuals(logit_m6, type = "deviance"), # Deviance residuals
                  Predictions = fitted(logit_m6),
                  x = data_3mi[[1]]$x,
                  y = data_3mi[[1]]$y)

RF2 <- RF2 |> filter(!is.na(Residuals))


Loc <- RF2[, c("x", "y")] #Spatial locations
Distances <- dist(Loc)

point.color = viridis::viridis(100,option = "F")
line.color = "gray10"
fill.color = viridis::viridis(4,option = "F",alpha = 0.95)[2]

#getting residuals
residuals <- RF2$Residuals
residuals.df <- as.data.frame(residuals)

 #normality scores of the residuals
residuals.qq <- qqnorm(residuals, plot.it=FALSE) %>%
  as.data.frame()

#qqplot
p1 <- ggplot2::ggplot(data = residuals.qq) +
  ggplot2::geom_point(
    data = residuals.qq,
    ggplot2::aes(
      x = x,
      y = y,
      color = y
    )
  ) +
  ggplot2::scale_color_gradientn(colors = point.color) +
  ggplot2::stat_qq_line(
    ggplot2::aes(sample = residuals),
    col = line.color,
    linetype = "dashed"
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::ylab("Residuals") +
  ggplot2::xlab("Theoretical")

#histogram
p2 <- ggplot2::ggplot(data = residuals.df) +
  ggplot2::aes(
    x = residuals
    ) +
  ggplot2::geom_histogram(
    # binwidth = bw,
    color = NA,
    fill = fill.color
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::geom_vline(
    xintercept = median(residuals),
    col = line.color,
    linetype = "dashed"
  ) +
  ggplot2::ylab("Count") +
  ggplot2::xlab("Residuals")

  #residuals vs predictions
p3 <- ggplot2::ggplot(RF2) +
    ggplot2::aes(
      x = Predictions,
      y = Residuals,
      color = Residuals
    ) +
    ggplot2::geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = line.color
    ) +
    ggplot2::geom_point() +
    ggplot2::scale_color_gradientn(colors = point.color) +
    ggplot2::theme_bw() +
    ggplot2::ggtitle("Residuals vs. predictions") +
    ggplot2::theme(
      plot.title = ggplot2::element_text(hjust = 0.5),
      legend.position = "none"
      )

variogram_data <- RF2
coordinates(variogram_data) <- ~x+y
proj4string(variogram_data)= sf::st_crs(3578)[[2]]
variogram_res <- gstat::variogram(Residuals~1, variogram_data, cressie = TRUE, cutoff = 100000)

p5 <- ggplot() +
  geom_point(data = variogram_res, aes(x = dist, y = gamma)) +
  geom_smooth(data = variogram_res, aes(x = dist, y = gamma),col = "red") +
  xlab("Distance m") + ylab("Semi-variance") +
  theme_bw() +
  theme(text = element_text(size = 11),
        legend.position="none") +
  ylim(0, 1) 

#final plot
residuals_plot <- ((p1 + p2) / p3) / (p5) + patchwork::plot_annotation(
      theme = ggplot2::theme(plot.title = element_text(hjust = 0.5))
  )
```

```{r}
residuals_plot
```

```{r}
fit.variogram(variogram_res, vgm("Sph"))
```

```{r}
library(PsychHelperFunctions)
# Asses for OD
PsychHelperFunctions::overdisp_fun(logit_m6)
```


